---
title: "Vignette Title"
author: "József Varga"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{noaa - Getting and displaying Earthquake data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction
This document serves as the main overview of the noaa package, and will try 
to explain the hows and whys of the package content along with clear visual
examples.

# NOAA data

The  U.S. National Oceanographic and Atmospheric Administration (NOAA) publish a database
on significant earthquakes around the world.
The dataset contains information about 5,933 earthquakes over an approximately 4,000 year time span.

The datasset is avialable at 
[https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1](NOAA Link) . An actual state of the dataset can be obtained at the "Download" link.


## Package snapshot
In order to support easy access to the dataset the package contains a snapshot of the avialable data.
To use the data it should be loaded:

```{r}
data(earthquakes)
```

## Use the latest data

It is useful to use the latest downloaded data from the NOAA site. `eq_read_data()` is the function to use:

```{r}
eq <- eq_read_data(filename="data-raw/signif.txt.bz2" )
```

## Field description


The detailed description of the dataset is supported by NOAA at the "Event Variable Definitions" link.

The documentation also included in the package: `man/earthquakes.Rd`


# Data cleaning

The raw NOAA data is not ideal for R processing. There are two helper functions to clean the data.

## Datetime field insertion

NOAA dataset contains separate fields for the date - time: 

## Location clean

# Displaying a timeline

# Displaying an interactive map

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
