---
title: "Vignette Title"
author: "Jozsef Varga"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{noaa - Getting and displaying Earthquake data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
library(noaa)
knitr::opts_chunk$set(fig.dpi = 96)
```

# Introduction
This document serves as the main overview of the noaa package, and will try 
to explain the hows and whys of the package content along with clear visual
examples.

# NOAA data

The  U.S. National Oceanographic and Atmospheric Administration (NOAA) publish a database
on significant earthquakes around the world.
The dataset contains information about 5,933 earthquakes over an approximately 4,000 year time span.

The datasset is avialable at 
[NOAA Link](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1) . An actual state of the dataset can be obtained at the "Download" link.


## Package snapshot
In order to support easy access to the dataset the package contains a snapshot of the avialable data.
To use the data it should be loaded:

```{r}
data(earthquakes)
```


## Use the latest data

It is useful to use the latest downloaded data from the NOAA site. `eq_read_data()` is the function to use:

```{r}
eq <- eq_read_data(filename="../data-raw/signif.txt.bz2" )
```

## Field description


The detailed description of the dataset is supported by NOAA at the "Event Variable Definitions" link.

The documentation also included in the package: `man/earthquakes.Rd`, use ` ?earthquakes` to see it.


# Data cleaning

The raw NOAA data is not ideal for R processing. There are two helper functions to clean the data.

## Datetime field insertion

NOAA dataset contains separate fields for the date - time: 'YEAR','MONTH','DAY','HOUR','MINUTE','SECOND'.
`eq_clean_data()` create a new R datetime field `date` and deletes the separate fields.

To clean the built in earthquakes data you should run : 
```{r}
data(earthquakes)
eq <- eq_clean_data(earthquakes)
head(eq$date)
```

## Location clean

NOAA dataset has a "LOCATION" filed containing all caps text wth country information added.
To display this field a simple transformation is done by `eq_location_clean`.  The function strips out 
the "COUNTRY" part and formats the string as "Title Case".

```{r}
data(earthquakes)
# NOAA version
head(earthquakes$LOCATION_NAME)

eq <- eq_location_clean(earthquakes)
# Cleaned version
head(eq$LOCATION_NAME)
```

# Displaying a timeline

The package contains a ggplot2 geom to visualize the times at which earthquakes occur within certain countries.

The `geom_timeline()` main use to display a timeline of events:

The function `geom_timeline_label()` adds text directly to the timeline plot.
The param `n_max` defines the number of the max size events to display.


```{r fig1, fig.height = 3, fig.width = 8}
library(magrittr)
library(ggplot2)
library(lubridate)
library(dplyr)
data(earthquakes)
eqdta <- eq_location_clean(eq_clean_data(earthquakes)) %>%
filter(date > lubridate::ymd("20000101") &
(COUNTRY == "USA" | COUNTRY == 'CHINA')) %>%
group_by(COUNTRY)

# Draw the geom
ggplot (data = eqdta,
aes(
x = date,
y = COUNTRY,
size = FOCAL_DEPTH,
label = LOCATION_NAME,
colour = EQ_PRIMARY
)) +
geom_timeline() +
geom_timeline_label(n_max = 6)
```

 
# Displaying an interactive map

